Evaluation

### 2018-02-02 evaluation.Rmd

species_of_interest = c('BA','DM','DO','PP','OT','NA','total')
species_names = portalr::FullPath('PortalData/Rodents/Portal_rodent_species.csv', '~') %>%
  read.csv(stringsAsFactors=FALSE) %>%
  select(species = speciescode, full_species_name = scientificname)

#Fix neotoma, and add a total entry
species_names$species[species_names$full_species_name=='Neotoma albigula'] <- "NA"
species_names = species_names %>%
  mutate(species = ifelse(full_species_name=='Neotoma albigula', 'NA', species)) %>%
  add_row(species='total', full_species_name='Total Rodents')

@@

new_moon_file = portalr::FullPath('PortalData/Rodents/moon_dates.csv', '~')
new_moons = read.csv(new_moon_file)
most_recent_newmoon = max(new_moons$newmoonnumber)
earliest_newmoon_to_graph = most_recent_newmoon - prior_new_moons

species_abundance = portalr::abundance(shape='flat', level='Site') %>%
  rename(actual = abundance) %>%
  mutate(level='All',currency='abundance') %>%
  left_join(new_moons, by='period') %>%
  filter(species %in% species_of_interest)

@@

#Get the all the forecasts made during observation period
forecast_data = compile_forecasts(use_hindcasts = FALSE) %>%
  filter(newmoonnumber %in% earliest_newmoon_to_graph:most_recent_newmoon)

@@

# Keep 1 forecast per unique initial_newmoon, when there are multiple forecasts
# using the same initial_newmoon, use the first one.
forecast_dates_to_keep = forecast_data %>%
  select(date, initial_newmoon) %>%
  distinct() %>%
  group_by(initial_newmoon) %>%
  filter(date == min(date)) %>%
  ungroup() %>%
  mutate(keep='yes')

@@

####################################
# Setup a nice palette for when prior issue dates become greater than 9
getPalette = colorRampPalette(RColorBrewer::brewer.pal(9, "Set1"))
#Randomize pallette so nearby regions are not similar colors
set.seed(2)
large_color_palette = sample(getPalette(40))
#####################################
ggplot(forecast_errors, aes(x=censusdate, y=rmse, group=as.character(date), color=as.character(date))) +
  geom_point(size=4)+
  geom_line(size=1.5) +
  # label individual issue dates. This is easier to look at  if there 
  # gets to be a lot of them. Must install ggrepel and remove the legend if this is turned on.
  #geom_label_repel(data=forecast_date_labels, aes(label=label_text, y=y_position),
  #                 fill='grey90',size=4,nudge_y = 40, force=5,
  #                 segment.alpha = 0.3, segment.size = 0.8) +
  scale_color_brewer(palette='Dark2') +
  #scale_color_manual(values=large_color_palette) +
  facet_grid(species~model) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=0.9, debug = FALSE),
        axis.text = element_text(size=12),
        axis.title = element_text(size=20),
        strip.text = element_text(size=12),
        legend.text = element_text(size=15),
        legend.title = element_text(size=18)) +
  labs(x='Sample Date',y='RMSE', color='Forecast Issue\nDate')

@@

species_abundance = portalr::abundance(shape='flat', level='Site') %>%
  rename(actual = abundance) %>%
  mutate(level='All',currency='abundance') %>%
  left_join(new_moons, by='period') %>%
  filter(species %in% species_of_interest)

>>>>

species_of_interest = c('BA','DM','DO','PP','OT','NA','total')
species_names = portalr::FullPath('PortalData/Rodents/Portal_rodent_species.csv', '~') %>%
  read.csv(stringsAsFactors=FALSE, na.strings = '') %>%
  select(species = speciescode, full_species_name = scientificname)

#add a total entry
species_names = species_names %>%
  add_row(species='total', full_species_name='Total Rodents')

# Setup rodent observations
species_abundance = portalr::abundance(shape='flat', level='Site') %>%
  rename(actual = abundance) %>%
  mutate(level='All',currency='abundance') %>%
  left_join(new_moons, by='period') %>%
  filter(species %in% species_of_interest)

@@

most_recent_newmoon = max(new_moons$newmoonnumber)
earliest_newmoon_to_graph = most_recent_newmoon - prior_new_moons

@@

#Get all the recent forecasts
forecast_data = compile_forecasts(use_hindcasts = FALSE) %>%
  filter(newmoonnumber %in% earliest_newmoon_to_graph:most_recent_newmoon)

@@

# Keep 1 forecast per unique initial_newmoon. When there are multiple forecasts
# using the same initial_newmoon, use the first one.
forecast_dates_to_keep = forecast_data %>%
  select(date, initial_newmoon) %>%
  distinct() %>%
  group_by(initial_newmoon) %>%
  filter(date == min(date)) %>%
  ungroup() %>%
  mutate(keep='yes')

@@

####################################
# Setup a nice palette for when prior issue dates become greater than 9
getPalette = colorRampPalette(RColorBrewer::brewer.pal(9, "Set1"))
set.seed(2)
large_color_palette = sample(getPalette(40))
#####################################
ggplot(forecast_errors, aes(x=censusdate, y=rmse, group=as.character(date), color=as.character(date))) +
  geom_point(size=4)+
  geom_line(size=1.5) +
  # label individual issue dates. This is easier to look at if there 
  # gets to be a lot of them. Must install ggrepel and remove the legend if this is turned on.
  #geom_label_repel(data=forecast_date_labels, aes(label=label_text, y=y_position),
  #                 fill='grey90',size=4,nudge_y = 40, force=5,
  #                 segment.alpha = 0.3, segment.size = 0.8) +
  scale_color_brewer(palette='Dark2') +
  #scale_color_manual(values=large_color_palette) +
  facet_grid(species~model) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=0.9, debug = FALSE),
        axis.text = element_text(size=12),
        axis.title = element_text(size=20),
        strip.text = element_text(size=12),
        legend.text = element_text(size=15),
        legend.title = element_text(size=18)) +
  labs(x='Sample Date',y='RMSE', color='Forecast Issue\nDate')

@@
